---
title: "Homework 2 Report"
output: html_document
---


## Name

Mason Reed

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.
maacs = read.csv('maacs.csv.gz')
library(ggplot2)
library(tidyverse)
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}
## Write your function here

fev1_age_positive <- function(df) {

  # fit linear regression of FEV1 on age
  fev1_m <- lm(fev1 ~ age, data = df)
  
  # extract the slope for age
  slope <- coef(fev1_m)["age"]
  
  # return TRUE if slope is positive, FALSE otherwise
  return(slope > 0)
}

  



fev1_age_positive(maacs)

```


## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 

```{r}

fev1_m <- lm(fev1 ~ age, data = maacs)
coef(fev1_m)
```


Write your data analysis statement interpreting the regression model here:

The regression model suggests that fev1 increases with age by 0.171 for an increase of 1 year in age.

## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results. 

NOTE: 

* At least one supporting premise should use a plot.

* Do not use the code you write in Part 1 as a supporting premise

```{r}
## Add your code here

# Premise 1 Scatterplot
ggplot(maacs, aes(x = age, y = fev1)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +  
  labs(
    title = "Scatterplot of FEV1 vs Age (Supporting Premise 1)",
    x = "Age (years)",
    y = "FEV1"
  ) +
  theme_minimal()

fev1_summary = summary(fev1_m)


  
  # extract slope estimate and p-value and correlation for age
  slope <- coef(fev1_m)["age"]
  p_val = coef(fev1_summary)["age", "Pr(>|t|)"]
  correlation = sqrt(fev1_summary$r.squared)
  
  # Slope P-Value Correlation Table
   results <- data.frame(
    Slope_Estimate = slope,
    P_Value = p_val,
    Correlation = correlation,
    Slope_Positive = slope > 0,
    Significant_Positive = (p_val < 0.05 & slope > 0)
  )
  results


```


Write the three supporting premise statements here:

1. The scatterplot shows fev1 generally increasing as the individuals are older.

2. The correlation of `r round(correlation,3)` suggests a strong, positive relationship which supports the positive slope value of 0.171.

3. The positive slope in combination with the small p-value for the slope supports the positive relationship between fev1 and age.


## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
## Function for supporting premise statement 1
scatter <- function(dataframe) {
ggplot(dataframe, aes(x = age, y = fev1)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +  
  labs(
    title = "Scatterplot of FEV1 vs Age (Premise 1 Test)",
    x = "Age (years)",
    y = "FEV1"
  ) +
  theme_minimal()
}

```


```{r}
## Function for supporting premise statement 2
check_correlation_plus <- function(df){
  fev1_lm = lm(fev1~age,df)
  fev1_summary = summary(fev1_lm)
  correlation = sqrt(fev1_summary$r.squared)
  return (correlation > 0)
}
```


```{r}
## Function for supporting premise statement 3
premise3_pval_test <- function(df) {
  
  fit <- lm(fev1 ~ age, data = df)
  
  # get summary of the regression
  fit_summary <- summary(fit)
  
  # extract slope estimate and p-value for age
  slope <- coef(fit)["age"]
  p_val <- coef(fit_summary)["age", "Pr(>|t|)"]
  
  return (slope > 0 & p_val < 0.01)
 
}
```

Execute each of your function and show that the produce the expected output.
```{r}
# Premise 1 Test : Scatterplot
scatter(maacs)

# Premise 2 Test: Correlation Coef
check_correlation_plus(maacs)

# Premise 3 Test: P-val and Positive Slope
premise3_pval_test(maacs)


```



## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children". 

Alternative: FEV1 Values are the same for older children relative to younger children.

Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.


